---
title: "project 2"
author: "Sichun Li, Qiqi Liu, Tong Niu, Yunqing Yu, Xiao Yang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r}
library("dummies")
library("AER")
library("scatterplot3d")
library("rgl")
library("data.table")
library("mlogit")
library("gmnl")
rm(list = ls());
setwd("~/Desktop/Pricing Analytics/project/project 2")
```


```{r load data}
data=fread("kiwi_bubbles_P2.csv",stringsAsFactors = F)
#Data cleaning - drop periods with price=99 (stockout).
data=data[!(data$price.KB==99),]
data=data[!(data$price.KR==99),]
data=data[!(data$price.MB==99),]
```

3 Logit model without segmentation
```{r Q3:Logit model without segmentation}
#Multinomial logit
#Product-line pricing
#Solve a profit maximization problem over two products

#Write choice probability for both KB and KR as a function
#Notational change - use "para" to represent all parameter inputs,
#instead of separately defining them as "beta0KB,beta0KR,beta0MB,beta1".

demand=function(priceKB,priceKR,priceMB,para){
  probKB=exp(para[1]+para[4]*priceKB)/(1+exp(para[1]+para[4]*priceKB)+exp(para[2]+para[4]*priceKR)+exp(para[3]+para[4]*priceMB))
  probKR=exp(para[2]+para[4]*priceKR)/(1+exp(para[1]+para[4]*priceKB)+exp(para[2]+para[4]*priceKR)+exp(para[3]+para[4]*priceMB))
  return(cbind(probKB,probKR))
}

#Write profit as a function of prices we set and model parameters
profit=function(priceKB,priceKR,priceMB,para){
  profitKB=demand(priceKB,priceKR,priceMB,para)[,1]*(priceKB-uc)
  profitKR=demand(priceKB,priceKR,priceMB,para)[,2]*(priceKR-uc)
  return(cbind(profitKB,profitKR))
}
#Unit cost
uc=0.5;

###Estimation of multinomial logit model
#Now columns 4 through 7 contains "Price.something" info.
mlogitdata=mlogit.data(data,id="id",varying=4:7,choice="choice",shape="wide")
#Run MLE.
mle= gmnl(choice ~  price, data = mlogitdata)
summary(mle)
coef=mle$coefficients

#Set parameter
#The first element of "para" is beta0KB,beta0KR,beta0MB,beta1"
para=c(coef[1],coef[2],coef[3],coef[4])

###calculate own- and cross-price elasticities(evaluated at the average prices observed in the data)
meanPrice<-c(mean(data$price.KB),mean(data$price.KR),mean(data$price.MB))
demandForAll=function(priceKB,priceKR,priceMB,para){
  probKB=exp(para[1]+para[4]*priceKB)/(1+exp(para[1]+para[4]*priceKB)+exp(para[2]+para[4]*priceKR)+exp(para[3]+para[4]*priceMB))
  probKR=exp(para[2]+para[4]*priceKR)/(1+exp(para[1]+para[4]*priceKB)+exp(para[2]+para[4]*priceKR)+exp(para[3]+para[4]*priceMB))
  probMB=exp(para[3]+para[4]*priceMB)/(1+exp(para[1]+para[4]*priceKB)+exp(para[2]+para[4]*priceKR)+exp(para[3]+para[4]*priceMB))
  return(cbind(probKB,probKR,probMB))
}
prob<-demandForAll(meanPrice[1],meanPrice[2],meanPrice[3],para)

ownElasticity=function(beta1,price,prob){
  ownElasticity=-beta1*price*(1-prob)
  return(ownElasticity)
}
crossElasticity=function(beta1,price,prob){
  crossElasticity=-beta1*price*prob
  return(crossElasticity)
}

#built a matrix for both own- and cross- ealsticity 

ElastMatrix<-data.frame(matrix(ncol = 3, nrow = 3))
colnames(ElastMatrix)<- c("KB",'KR','MB')
rownames(ElastMatrix)<- c("KB",'KR','MB')

for(i in 1:3){
  for(j in 1:3){#Products in column change prices and then influence products in rows
    if (rownames(ElastMatrix)[i]==colnames(ElastMatrix)[j]){
      ElastMatrix[i,j]=ownElasticity(para[4],meanPrice[i],prob[i])}
    else{ElastMatrix[i,j]=crossElasticity(para[4],meanPrice[j],prob[j])}
  }
}


###calculate profit
#"demand" function represents each individual consumer's choice probability.
#In order to calculate profit, we multiply the "demand" by the number of consumers.

#Choose space of prices to search for the optimal price over
aux=seq(0.88,1.47,0.01)
#Because we search over two dimensions, create complete combination 
#of the two prices
pricespace=expand.grid(aux,aux)
colnames(pricespace)=c('priceKB','priceKR')
#Compute profit at each realization of this price space.
#write for-loop, take one realization of [P^KB,P^KR] pair and evaluate
#profit at that realization.
profitmat=matrix(0L,nrow(pricespace),1)
for (i in 1:nrow(pricespace)){
    profitmat[i]=sum(profit(pricespace[i,1],pricespace[i,2],1.43,para))  
}
expectedProfit=1000*max(profitmat[,1])
optimalPrices=pricespace[which.max(profitmat[,1]),]
optimalPrices
#Both optimal KB price and optimal KR are 1.16.
```

4 Logit model with segmentation
```{r Q4:Logit model with segmentation}
library(cluster)
library(fpc)
library(factoextra)
library(gridExtra)
library(conjoint)
demo=fread("demo_P2.csv",stringsAsFactors = F)
#Number of individuals
N = length(unique(data$id))

#Clustering
clustTest = function(toClust,print=TRUE,scale=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100){
    if(scale){ toClust = scale(toClust);}
    set.seed(seed);   # set random number seed before doing cluster analysis
    wss <- (nrow(toClust)-1)*sum(apply(toClust,2,var))
    for (i in 2:maxClusts) wss[i] <- sum(kmeans(toClust,centers=i,nstart=nstart,iter.max=iter.max)$withinss)
    ##gpw essentially does the following plot using wss above. 
    #plot(1:maxClusts, wss, type="b", xlab="Number of Clusters",ylab="Within groups sum of squares")
    gpw = fviz_nbclust(toClust,kmeans,method="wss",iter.max=iter.max,nstart=nstart,k.max=maxClusts) #alternative way to get wss elbow chart.
    pm1 = pamk(toClust,scaling=TRUE)
    ## pm1$nc indicates the optimal number of clusters based on 
    ## lowest average silhoutte score (a measure of quality of clustering)
    #alternative way that presents it visually as well.
    gps = fviz_nbclust(toClust,kmeans,method="silhouette",iter.max=iter.max,nstart=nstart,k.max=maxClusts) 
    if(print){
        grid.arrange(gpw,gps, nrow = 1)
    }
    list(wss=wss,pm1=pm1$nc,gpw=gpw,gps=gps)
}
##Runs a set of clusters as kmeans
##Arguments:
##  toClust, data.frame with data to cluster
##  nClusts, vector of number of clusters, each run as separate kmeans 
##  ... some additional arguments to be passed to clusters
##Return:
##  list of 
##    kms, kmeans cluster output with length of nClusts
##    ps, list of plots of the clusters against first 2 principle components
runClusts = function(toClust,nClusts,print=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100){
    kms=list(); ps=list();
    for(i in 1:length(nClusts)){
        kms[[i]] = kmeans(toClust,nClusts[i],iter.max = iter.max, nstart=nstart)
        ps[[i]] = fviz_cluster(kms[[i]], geom = "point", data = toClust) + ggtitle(paste("k =",nClusts[i]))
        
    }
    library(gridExtra)
    if(print){
        tmp = marrangeGrob(ps, nrow = 2,ncol=2)
        print(tmp)
    }
    list(kms=kms,ps=ps)
}
##Plots a kmeans cluster as three plot report
##  pie chart with membership percentages
##  ellipse plot that indicates cluster definitions against principle components
##  barplot of the cluster means
plotClust = function(km,toClust,discPlot=FALSE){
    nc = length(km$size)
    if(discPlot){par(mfrow=c(2,2))}
    else {par(mfrow=c(3,1))}
    percsize = paste(1:nc," = ",format(km$size/sum(km$size)*100,digits=2),"%",sep="")
    pie(km$size,labels=percsize,col=1:nc)
    
    clusplot(toClust, km$cluster, color=TRUE, shade=TRUE,
             labels=2, lines=0,col.clus=1:nc); #plot clusters against principal components
    
    if(discPlot){
        plotcluster(toClust, km$cluster,col=km$cluster); #plot against discriminant functions ()
    }
    rng = range(km$centers)
    dist = rng[2]-rng[1]
    locs = km$centers+.05*dist*ifelse(km$centers>0,1,-1)
    bm = barplot(km$centers,beside=TRUE,col=1:nc,main="Cluster Means",ylim=rng+dist*c(-.1,.1))
    text(bm,locs,formatC(km$centers,format="f",digits=1))
}

checks = clustTest(demo)
clusts=runClusts(demo,2:5)
for(i in 1:4) {plotClust(clusts[[1]][[i]],demo)}
   
KmeansCluster = function(n){
  set.seed(123)
  demo_cluster = kmeans(x=demo[, 2:18], centers = n, nstart = 1000)
  cluster_id = data.frame(id = demo$id)
  cluster_id$cluster = demo_cluster$cluster
  datafull = merge(data, cluster_id, by = "id", all.x = T)
  datafull$cluster[is.na(datafull$cluster)] = n+1
  N = length(unique(data$id))
  seg.share = c( table(demo_cluster$cluster),N - sum(table(demo_cluster$cluster))) / N 

# just store the coefficients (you can store many other things)
  coef.est = data.frame(segment = 1:(n+1), intercept.KB = NA, intercept.KR = NA, 
                      intercept.MB = NA, price.coef = NA) 
    #Write a for-loop. 
  for (seg in 1:(n+1)) {
  # During each loop, pick subset of data of consumers from each segment.
      data.sub = subset(datafull, cluster %in% seg)
      mlogitdata=mlogit.data(data.sub,id="id",varying=4:7,choice="choice",shape="wide")
  
  #Run MLE.
      mle= gmnl(choice ~  price, data = mlogitdata)
  #Store the outcome in the coef.est matrix.
      coef.est[seg, 2:5] = mle$coefficients
        
}
      return(list(seg.share,datafull,coef.est))
}

KmeansCluster(2)  #change 0
KmeansCluster(3)  #change 11.2462%
KmeansCluster(4)  #change 10.6383%
KmeansCluster(5)  #change 10.6383%
KmeansCluster(6)  #change 0
KmeansCluster(7)  #change 11.550152%
KmeansCluster(8)  #change 11.550152%
KmeansCluster(9)  #change 11.550152%
KmeansCluster(10) #5,8 change 0.11550152+0.07294833=0.1884499 ######seems the best
KmeansCluster(11) #1,3 change 0.07294833+0.09422492=0.1671732
KmeansCluster(12) #6   change 0.07294833
KmeansCluster(13) #change 0
KmeansCluster(14) #6 change 0.08814590
KmeansCluster(15) #6 change 0.08814590
###So, when centers=10, segment=11, after lauching KB, people's change from MB to KB ......
seg.share = KmeansCluster(10)[[1]]
coef.est = KmeansCluster(10)[[3]]
```


4.1 Please see our report.

4.2.1 what are the (product-level, aggregated across segments) own- and cross- elasticities among these products?
```{r}
#Calculate elasticity
prSeg=function(priceKB,priceKR,priceMB) {
  Pr1=demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[1,2:5]))
  Pr2=demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[2,2:5]))
  Pr3=demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[3,2:5]))
  Pr4=demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[4,2:5]))
  Pr5=demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[5,2:5]))
  Pr6=demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[6,2:5]))
  Pr7=demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[7,2:5]))
  Pr8=demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[8,2:5]))
  Pr9=demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[9,2:5]))
  Pr10=demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[10,2:5]))
  Pr11=demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[11,2:5]))
  return(list(Pr1,Pr2,Pr3,Pr4,Pr5,Pr6,Pr7,Pr8,Pr9,Pr10,Pr11))
}# each c(Pr_KB,Pr_KR,Pr_MB)

agg_choice=function(priceKB,priceKR,priceMB) {
  agg_choice=seg.share[1]*demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[1,2:5]))+
    seg.share[2]*demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[2,2:5]))+
    seg.share[3]*demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[3,2:5]))+
    seg.share[4]*demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[4,2:5]))+
    seg.share[5]*demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[5,2:5]))+
    seg.share[6]*demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[6,2:5]))+
    seg.share[7]*demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[7,2:5]))+
    seg.share[8]*demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[8,2:5]))+
    seg.share[9]*demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[9,2:5]))+
    seg.share[10]*demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[10,2:5]))+
    seg.share[11]*demandForAll(priceKB,priceKR,priceMB,as.numeric(coef.est[11,2:5]))
  return(agg_choice)
}# aggregate c(Pr_KB,Pr_KR,Pr_MB) R,Pr_MB) 
segPr=prSeg(meanPrice[1],meanPrice[2],meanPrice[3])
segPr <- data.frame(matrix(unlist(segPr), nrow=length(segPr), byrow=T))
segAggPr=agg_choice(meanPrice[1],meanPrice[2],meanPrice[3])
 
# own-price elasticities
segOwnElasticity=function(price,Pr,segShare,beta0,segPr){
  SegOwnElasticity=-price*sum(segShare*beta0*segPr*(1-segPr))/Pr
  return(SegOwnElasticity)
}
# cross-price elasticities j's price to i
segCrossElasticity=function(price,Pr,segShare,beta1,segPr,i,j){
  SegCrossElasticity=-price/Pr*sum(segShare*beta1*segPr[[i]]*segPr[[j]])
  return(SegCrossElasticity)
}

#built a matrix for both own- and cross- ealsticity 
segElastMatrix<-data.frame(matrix(ncol = 3, nrow = 3))
colnames(segElastMatrix)<- c("KB",'KR','MB')
rownames(segElastMatrix)<- c("KB",'KR','MB')

for(i in 1:3){
  for(j in 1:3){#Products in column change prices and then influence products in rows
    if (rownames(segElastMatrix)[i]==colnames(segElastMatrix)[j]){
      segElastMatrix[i,j]=segOwnElasticity(meanPrice[i],segAggPr[i],seg.share,coef.est[,5],segPr[[i]])
    }else{segElastMatrix[i,j]=segCrossElasticity(meanPrice[j],segAggPr[i],seg.share,coef.est[,5],segPr,i,j)}
  }
}
```
4.2.2 How does the underlying customer segmentation explain the substitution pattern you see in the elasticity?
From the substitution pattern and underlying segmentation, where (i.e. which segment(s)) should Kiwi Bubbles be positioned?

Please see our report.



4.2.3 
```{r}
#If not lauching KB
#"newpara" is beta0KB,beta0KR,beta0MB,beta1
demandPrevious=function(priceKR,priceMB,newpara){
  probKR=exp(newpara[2]+newpara[4]*priceKR)/(1+exp(newpara[2]+newpara[4]*priceKR)+exp(newpara[3]+newpara[4]*priceMB))
  probMB=exp(para[3]+para[4]*priceMB)/(1+exp(para[2]+para[4]*priceKR)+exp(para[3]+para[4]*priceMB))
  return(cbind(probKR,probMB))
}

agg_choicePrevious=function(priceKR,priceMB) {
  agg_choice=seg.share[1]*demandPrevious(priceKR,priceMB,as.numeric(coef.est[1,2:5]))+
    seg.share[2]*demandPrevious(priceKR,priceMB,as.numeric(coef.est[2,2:5]))+
    seg.share[3]*demandPrevious(priceKR,priceMB,as.numeric(coef.est[3,2:5]))+
    seg.share[4]*demandPrevious(priceKR,priceMB,as.numeric(coef.est[4,2:5]))+
    seg.share[5]*demandPrevious(priceKR,priceMB,as.numeric(coef.est[5,2:5]))+
    seg.share[6]*demandPrevious(priceKR,priceMB,as.numeric(coef.est[6,2:5]))+
    seg.share[7]*demandPrevious(priceKR,priceMB,as.numeric(coef.est[7,2:5]))+
    seg.share[8]*demandPrevious(priceKR,priceMB,as.numeric(coef.est[8,2:5]))+
    seg.share[9]*demandPrevious(priceKR,priceMB,as.numeric(coef.est[9,2:5]))+
    seg.share[10]*demandPrevious(priceKR,priceMB,as.numeric(coef.est[10,2:5]))+
    seg.share[11]*demandPrevious(priceKR,priceMB,as.numeric(coef.est[11,2:5]))
  return(agg_choice)
}

uc=0.5
pricespaceForKR=seq(0.88,1.43,0.01)
profit_previous=1000*(agg_choicePrevious(pricespaceForKR,1.43)[,1])*(pricespaceForKR-uc)
max(profit_previous)
priceKR_previous=pricespaceForKR[which.max(profit_previous)]
priceKR_previous
#if we do not launch KB, optimal price is $1.06, best profit is 285.592

#If lauching KB
#The first element of "newpara" is beta0KB,beta0KR,beta0MB,beta1"
ProfitAfter=function(priceKB,priceKR,priceMB){
  profitKB=agg_choice(priceKB,priceKR,priceMB)[,1]*(priceKB-0.5)*1000
  profitKR=agg_choice(priceKB,priceKR,priceMB)[,2]*(priceKR-0.5)*1000
  profitMB=agg_choice(priceKB,priceKR,priceMB)[,3]*(priceMB-0.5)*1000
  return(cbind(profitKB,profitKR,profitMB))
}
#Choose space of prices to search for the optimal price over
aux=seq(0.88,1.47,0.01)
#Because we search over two dimensions, create complete combination 
#of the two prices
pricespace=expand.grid(aux,aux)
#At each iteration of the loop, I take one realization of [P^KB,P^KR] pair and evaluate
#profit at that realization.
profitmat=matrix(nrow(pricespace),1)
for (i in 1:nrow(pricespace)){
    profitmat[i]=sum(ProfitAfter(pricespace[i,1],pricespace[i,2],1.43)[,c(1,2)])
}
priceKB_seg = pricespace[profitmat==max(profitmat),][,1];
priceKB_seg #1.13
priceKR_seg = pricespace[profitmat==max(profitmat),][,2];
priceKR_seg #1.2
profit_after = max(profitmat); 
profit_after #395.6119

#MB change, before launch KB
profitMB_previous=1000*agg_choicePrevious(priceKR_previous,1.43)[,2]*(1.43-uc)
profitMB_previous #105.6955
profitMB_after=ProfitAfter(priceKB_seg,priceKR_seg,1.43)[,3]
profitMB_after #86.57259
```




5 Understanding strategic responses
```{r Q5:Understanding strategic responses}
# price war

#As Mango, I need to react to KB and KR’s new prices.
KB1=priceKB_seg 
KR1=priceKR_seg
uc=0.5
pricespace1=seq(0,2,0.01)
profit1=1000*agg_choice(KB1,KR1,pricespace1)[,3]*(pricespace1-uc)
max(profit1)
MB1=pricespace1[profit1==max(profit1)];
MB1 #0.95


#As Kiwi, I need to react to MB’s new price.
aux2=seq(0.8,2,0.01)
pricespace2=expand.grid(aux2,aux2)
profitmat=matrix(0L,nrow(pricespace2),1)
for (i in 1:nrow(pricespace2)){
    profitmat[i]=sum(ProfitAfter(pricespace2[i,1],pricespace2[i,2],MB1)[,1:2])
    }
KB2 = pricespace2[profitmat==max(profitmat),][,1];
KB2 #0.99
KR2 = pricespace2[profitmat==max(profitmat),][,2];
KR2 #1.1

#Then, as mango, I need to react to KB and KR's newer prices
profit=1000*agg_choice(KB2,KR2,pricespace1)[,3]*(pricespace1-uc)
MB2=pricespace1[profit==max(profit)];
MB2 #0.91

#Then, as Kiwi, I need to react to MB’s newer price.
profitmat=matrix(0L,nrow(pricespace2),1)
for (i in 1:nrow(pricespace2)){
    profitmat[i]=sum(ProfitAfter(pricespace2[i,1],pricespace2[i,2],MB2)[,c(1,2)])
    }

KB3 = pricespace2[profitmat==max(profitmat),][,1];KB3 #0.98
KR3 = pricespace2[profitmat==max(profitmat),][,2];KR3 #1.09

#Then,then, as mango, I need to react to KB and KR's newerer prices
profit=1000*agg_choice(KB3,KR3,pricespace1)[,3]*(pricespace1-uc)
MB4=pricespace1[profit==max(profit)];
MB4 #0.91

```